Nabi - A Preventative Solution Against JavaScript and Scriptless Tabnabbing Attacks

### 2015-05-06 paper/paper.tex
Added content to the introduction and intro background section, references.

Phishing attacks are always developing and evolving, and one particular type of attack attempts to trick the user well after they have initially visited the site.  Tabnabbing is a specific phishing technique that relies on the user’s trust of websites they have already viewed and considered safe. Some users often leave many tabs open to return to interesting websites at a later time, and this behavior makes those users the perfect target for this type of attack. When the user has left a tab, a malicious website can detect that the user is not currently viewing that tab and change the content of that page. This is done using JavaScript or the HTML meta tag with a timer to redirect the page at a time when the user is most likely not viewing it. The malicious site will typically attempt to change the site to appear as a login page for a familiar site such as Gmail or Facebook.  When the user returns to the tab, the changed page may request the user for their username and password, citing reasons such as an expired login session.  Because tabs are not expected to have changed while out-of-focus, and because users do not check URLs often, the site will seem legitimate, and the user will be less cautious because the tab was not newly created.

\begin{comment}

Phishing attacks are becoming more common as web use increases, and as users become more aware of phishing attacks, phishing techniques adapt to continue to be effective. Tabnabbing is one such new technique that has emerged recently and is effective even against users that have been trained to only give sensitive information to trusted websites.
\end{comment}

\section{Background}
\begin{comment}
Phishing attacks are not a new occurrence; they have been around since there has been personal data or credentials to steal. However, ``tabnabbing'' is a relatively recent addition to the list of phishing techniques, coined by Aza Raskin in 2010 \cite{Raskin}. Tabnabbing is a phishing technique that takes advantage of two commonly exhibited user behaviors for modern browser usage. The first is that users often open several tabs in a browser window and navigate between them without closing previous tabs. The second is the user tendency to leave browsers open for extended periods of time while away from their computers in order to access the pages more easily upon their return. 
\end{comment}

\subsection{TabShots}
\begin{comment}
One approach to combating tabnabbing is an extension known as TabShots, a program created by Philippe De Ryck, Nick Nikiforakis, Lieven Desmet, and Wouter Joosen \cite{TabShots}. TabShots continuously takes screenshots of tab and compares them to the previous ones and highlights any changes in red yellow or green depending on how many changes were made. When the user navigates back to a tab that has been altered in a tabnabbing attack, they will see the highlighted portions of the screen and receive a notification that the tab has been altered. 
Performance inpact on larger screens
Relies on user to make decision
\end{comment}

One approach to combating tabnabbing is a Google Chrome browser extension known as TabShots, which was created by Philippe De Ryck, Nick Nikiforakis, Lieven Desmet, and Wouter Joosen \cite{TabShots}.  To detect whether a tab has changed or not while not in focus, TabShots uses visual comparison of the tab's appearance before and after the change in the user's focus.  The extension records screenshots of the currently focused tab at regular intervals, using a simple Google Chrome API call and storing the screenshot as a data URL.  The program also keeps record of the tab's favicon, or the icon displayed in the tab's url and title space.  These two sets of images provide the basis for comparison when a tab regains focus.

@@

According to the testings of the creators, TabShots is said to provide large compatibility with most major sites on the internet at a seemingly acceptable processing speed.  The extensions was used on the top 1000 websites listed on Alexa.com, and the results showed that 78\% of these sites fall within the safe threshold of 5\% or less in changed blocks, meaning no compatability issues exist with that page.  About 19\% of these sites have moderate changes less than 40\%, which is the threshold for high amounts of changes, and 3\% report more than 40\% differences in changed blocks.  While the last two numbers may seem high in terms of false positives, the authors note that TabShots never interferes with the functionality of a page, and does provide a whitelist to providing overlays for these sites.  In terms of speed, testing showed that TabShots was able to prepare and calculate the image comparisions on a $1366 \times 768$ window within an average time of 284ms, which should provide little impact on browser operating time.

There are some problems with this application.  First, the performance testing of TabShots was done on a $1366 \times 768$ window.  Many users today have monitors with much larger resolutions such as $1920 \times 1080$ or $2560 \times 1440$, which provide two or four times as many pixels.  This means that TabShots will run considerably longer on these types of monitors and provided a much larger performance strain.  TabShots is designed to calculate the comparisons every time the user changes tab, which lead to very frequent occurrences of calculations depending on user behavior, and can further exaggerate TabShot's slower performance on higher resolution monitors.  Finally, the visual comparison property of TabShots leads to many occurrences of false positives that lead to the high numbers mentioned above.  Examples of non-malicious behavior that would trigger overlay occurrences would be videos, GIFs, image slideshows, overlays, and other dynamic advertisements.  The authors note that slowly images and files would flag large amounts of tile changes.  In addition, if an element was added or removed from the page, every other element on the page would shift, which would flag a major change in the picture comparison algorithm.  The uncertain performance of TabShots on large monitors or frequently changing tabs, as well as its high opportunities for false positives make TabShots a potentially unideal solution to tabnabbing. 
\subsection{TabSol}
\begin{comment}
Another defense against tabnabbing, TabSol created by Amandeep Singh and Somanath Tripathy \cite{TabSol}, compares the hash values of the web page to detect changes and notifies the user. 
\end{comment}

Another attempt to detect tabnabbing is TabSol, a browser plugin created by Amandeep Singh and Somanath Tripathy \cite{TabSol}.  TabSol uses a whitelist and hash computation to determine if a web page has changed while out of focus.  TabSol attempts to detect tabnabbing using less computations and storage than other programs such as TabShots.

The process of TabShots begins with a simple whitelisting check, to check whether the newly opening URL has already been verified as legitimate.  If so, the process execution is stopped.  Otherwise, TabSol will calculate the hash digest of the web page source to save the current state of the page, using the cryptographic hash function SHA-1.  When the user switches back to the old tab, TabSol will recompute the hash digest of the page, and compare it to the stored digest.  If they match, TabSol determines that the site is legitimate and adds the page into the domain whitelist mentioned above.  Else, the page is treated as suspicious, and TabSol attempts to find a login form on the changed page, as the authors consider its presence necessary in a phishing attack.  The authors chose to sue the login form detction method presented in CANTINA+.  If a login form is present, then the page is considered a phishing web page attempting tabnabbing, and appropriate actions can be taken.

The performance of this implementation is a strong feature of this browser extension.  On an average machine with a 2.53 GHz processor, 4GB RAM and Windows 7 (64-bit), TabSol takes on average 157ms to detect a Tabnabbing attack, if it exists.  The process only uses two hash computations, and the most time-consuming part is the login form detection, which only is used on a subset of tab changes where the hash computations do not match.  In addition, the hash values are the only things needing to be stored, which takes up much less memory than screenshots of the user's windows in the case of TabShots.  Finally, the authors make the case that the login form detection is a crucial component of TabSol that is lacking in other tabnabbing detection systems, and prevents the large amounts of false positives in TabSol tabnabbing detection.

@@

\subsection{NoTabNab}
\begin{comment}
Another extension, NoTabNab created by Seckin Anil Unlu and Kemal Bicakci \cite{NoTabNab}, notifies the user when the favicon or the tab name change as well as when a new layer is added to an inactive tab. 
\end{comment}

NoTabNab is a browser addon designed by Secking Anil Unlu and Kemal Bicakci.  Their main objective was to focus on the favicon, page title, and layout changes for each tab, as the authors state that these properties must change in a tabnabbing attack.  In particular, the program looks at important offset and margin values that are calculated for the page after the CSS style rules and properties are applied to each HTML element.  The authors argue that any tabnabbing site must have some influence on these values, and thus they can be used to detect if the site has changed in an attempt of a tabnabbing attack.

@@

The authors point out the advantages of their minimal information gathering methodology.  The primary benefits of only recording data about the topmost elements is the reduced work over gathering the necessary data points for every element, which results in less storage and processing to initially store the data, and less processing to compare the various data points for each element in the document.  In addition, the comparison of values will be resource friendly, as only the simple favicon, title, and important CSS values are compared for each page, as opposed to entire high resolution pictures or entire documents.

>>>>

Phishing attacks are always developing and evolving, and one particular type of attack attempts to trick the user well after they have initially visited the site.  Coined by Firefox developer Aza Raskin in 2010, Tabnabbing is a specific phishing technique that relies on the user’s trust of websites they have already viewed and considered safe~\cite{Raskin}. Some users often leave many tabs open to return to interesting websites at a later time, and this behavior makes those users the perfect target for this type of attack. When the user has left a tab, a malicious website can detect that the user is not currently viewing that tab and change the content of that page. This is done using JavaScript or the HTML meta tag with a timer to rearrange the page at a time when the user is most likely not viewing it. The malicious site will typically attempt to change the page content, tab title, and favicon to appear as a login page for a familiar site such as Gmail or Facebook.  When the user returns to the tab, the changed page may request the user for their username and password, citing reasons such as an expired login session.  Because tabs are not expected to have changed while out-of-focus, and because users do not check URLs often, the site will seem legitimate and trustworthy.

Phishing attacks are becoming more common as web use increases, and as users become more aware of phishing attacks, phishing techniques adapt to continue to be effective. Tabnabbing is one such new technique that is effective even against users that have been trained to only give sensitive information to trusted websites.  It also takes advantage of the fact that users are most cautious when they first open a site, not when they return to an existing tab.  Because this specific type of attack has only been identified relatively recently, there have been fewer proposed solutions to detect or prevent tabnabbing attacks compared to other phishing techniques.

\section{Background}
\begin{comment}
Phishing attacks are not a new occurrence; they have been around since there has been personal data or credentials to steal. However, ``tabnabbing'' is a relatively recent addition to the list of phishing techniques, coined by Aza Raskin in 2010~\cite{Raskin}. Tabnabbing is a phishing technique that takes advantage of two commonly exhibited user behaviors for modern browser usage. The first is that users often open several tabs in a browser window and navigate between them without closing previous tabs. The second is the user tendency to leave browsers open for extended periods of time while away from their computers in order to access the pages more easily upon their return. 
\end{comment}

There are some preventative measures one can take to lower the risks of tabnabbing attacks.  Aza Raskin recommends using the Firefox Account Manager to prevent unwarranted credential release~\cite{Raskin}, but this can be cleverly overcome by loading the legitimate site and placing an identical but malicious form over the original form~\cite{NoTabNab}.  Many users recommend the browser extension NoScript, which stops JavaScript on every tab unless the site is whitelisted~\cite{NoTabNab}, but this extension is only available on Firefox, and the blanketed restriction of JavaScript may seem too extensive for some users.  There have been some applications that have developed specifically to combat tabnabbing, which will be explained through the following research.

\subsection{TabShots}
\begin{comment}
One approach to stopping tabnabbing is an extension known as TabShots, a program created by Philippe De Ryck, Nick Nikiforakis, Lieven Desmet, and Wouter Joosen~\cite{TabShots}. TabShots continuously takes screenshots of tab and compares them to the previous ones and highlights any changes in red yellow or green depending on how many changes were made. When the user navigates back to a tab that has been altered in a tabnabbing attack, they will see the highlighted portions of the screen and receive a notification that the tab has been altered. 
Performance inpact on larger screens
Relies on user to make decision
\end{comment}

One approach to combating tabnabbing is a Google Chrome browser extension known as TabShots, which was created by Philippe De Ryck, Nick Nikiforakis, Lieven Desmet, and Wouter Joosen~\cite{TabShots}.  To detect whether a tab has changed or not while not in focus, TabShots uses visual comparison of the tab's appearance before and after the change in the user's focus.  The extension records screenshots of the currently focused tab at regular intervals, using a simple Google Chrome API call and storing the screenshot as a data URL.  The program also keeps record of the tab's favicon, or the icon displayed in the tab's url and title space.  These two sets of images provide the basis for comparison when a tab regains focus.

@@

According to the testings of the creators, TabShots is said to provide large compatibility with most major sites on the internet at a seemingly acceptable processing speed.  The extensions was used on the top 1000 websites listed on Alexa.com, and the results showed that 78\% of these sites fall within the safe threshold of 5\% or less in changed blocks, meaning no compatability issues exist with that page.  About 19\% of these sites have moderate changes less than 40\%, which is the threshold for high amounts of changes, and 3\% report more than 40\% differences in changed blocks.  While the last two numbers may seem high in terms of false positives, the authors note that TabShots never interferes with the functionality of a page, and does provide a whitelist to providing overlays for these sites.  In terms of speed, testing showed that TabShots was able to prepare and calculate the image comparisions on a $1366 \times 768$ window within an average time of 284ms, which should provide little impact on browser operating time~\cite{TabShots}.

There are some problems with this application.  First, the performance testing of TabShots was done on a $1366 \times 768$ window.  Many users today have monitors with much larger resolutions such as $1920 \times 1080$ or $2560 \times 1440$, which provide two or four times as many pixels.  This means that TabShots will run considerably longer on these types of monitors and provided a much larger performance strain.  TabShots is designed to calculate the comparisons every time the user changes tab, which lead to very frequent occurrences of calculations depending on user behavior, and can further exaggerate TabShot's slower performance on higher resolution monitors.  Finally, the visual comparison property of TabShots leads to many occurrences of false positives that lead to the high numbers mentioned above.  Examples of non-malicious behavior that would trigger overlay occurrences would be videos, GIFs, image slideshows, overlays, and other dynamic advertisements.  The authors note that slowly images and files would flag large amounts of tile changes.  In addition, if an element was added or removed from the page, every other element on the page would shift, which would flag a major change in the picture comparison algorithm.  The uncertain performance of TabShots on large monitors or frequently changing tabs, as well as its high opportunities for false positives make TabShots a potentially unideal solution to tabnabbing. 
\subsection{TabSol}
\begin{comment}
Another defense against tabnabbing, TabSol created by Amandeep Singh and Somanath Tripathy~\cite{TabSol}, compares the hash values of the web page to detect changes and notifies the user. 
\end{comment}

Another attempt to detect tabnabbing is TabSol, a browser plugin created by Amandeep Singh and Somanath Tripathy~\cite{TabSol}.  TabSol uses a whitelist and hash computation to determine if a web page has changed while out of focus.  TabSol attempts to detect tabnabbing using less computations and storage than other programs such as TabShots.

The process of TabShots begins with a simple whitelisting check, to check whether the newly opening URL has already been verified as legitimate.  If so, the process execution is stopped.  Otherwise, TabSol will calculate the hash digest of the web page source to save the current state of the page, using the cryptographic hash function SHA-1~\cite{TabShots}.  When the user switches back to the old tab, TabSol will recompute the hash digest of the page, and compare it to the stored digest.  If they match, TabSol determines that the site is legitimate and adds the page into the domain whitelist mentioned above.  Else, the page is treated as suspicious, and TabSol attempts to find a login form on the changed page, as the authors consider its presence necessary in a phishing attack.  The authors chose to sue the login form detction method presented in CANTINA+.  If a login form is present, then the page is considered a phishing web page attempting tabnabbing, and appropriate actions can be taken.

The performance of this implementation is a strong feature of this browser extension.  On an average machine with a 2.53 GHz processor, 4GB RAM and Windows 7 (64-bit), TabSol takes on average 157ms to detect a Tabnabbing attack, if it exists~\cite{TabShots}.  The process only uses two hash computations, and the most time-consuming part is the login form detection, which only is used on a subset of tab changes where the hash computations do not match.  In addition, the hash values are the only things needing to be stored, which takes up much less memory than screenshots of the user's windows in the case of TabShots.  Finally, the authors make the case that the login form detection is a crucial component of TabSol that is lacking in other tabnabbing detection systems, and prevents the large amounts of false positives in TabSol tabnabbing detection.

@@

\subsection{NoTabNab}
\begin{comment}
Another extension, NoTabNab created by Seckin Anil Unlu and Kemal Bicakci~\cite{NoTabNab}, notifies the user when the favicon or the tab name change as well as when a new layer is added to an inactive tab. 
\end{comment}

NoTabNab is a browser addon designed by Secking Anil Unlu and Kemal Bicakci~\cite{NoTabNab}.  Their main objective was to focus on the favicon, page title, and layout changes for each tab, as the authors state that these properties must change in a tabnabbing attack.  In particular, the program looks at important offset and margin values that are calculated for the page after the CSS style rules and properties are applied to each HTML element.  The authors argue that any tabnabbing site must have some influence on these values, and thus they can be used to detect if the site has changed in an attempt of a tabnabbing attack.

@@

The authors point out the advantages of their minimal information gathering methodology.  The primary benefits of only recording data about the topmost elements is the reduced work over gathering the necessary data points for every element, which results in less storage and processing to initially store the data, and less processing to compare the various data points for each element in the document~\cite{NoTabNab}.  In addition, the comparison of values will be resource friendly, as only the simple favicon, title, and important CSS values are compared for each page, as opposed to entire high resolution pictures or entire documents.
